On the eletrodynamic of moving bodies

### 2019-06-03 Article_AS.Rmd
Intro JR

### Context
Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of canadian forest are harvested each year and an additionnal loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain ressource constant, it is therefore essential to compensate the sampling and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better stuctural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use. \ 
Improvement programs by genetic selection are in place for the most important commercial species [@DRF2000; @Mullin2011] including white spruce (Picea glauca (Moench) Voss) which has a significant economic value because of its geographical distribution and abundance [@Eyre1980; @Abrahamson2015]. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@Nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly accross almost all forested regions which represents a wide range of soils types and a climatic conditions [@Canadian2009]. \

### Qualité du bois et ses composantes
During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of lossing ressource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood quality has mainly two components: biologic and economic. The biologic component is the resistance to mortality by vigor and the economic component is the value of the wood for end-use. With regard to the biological component, in has been demonstrated [@Hietz2013; @osazuwa2014] that trees with a more uniform wood densitiy within the stem are also the more resistant to mortality. For example a uniform density gives a greater mechanical resistance to the wind and therefore a lower probability of mortality during a climatic event such as a storm. With regard to the economical context, several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. Indeed this trait is important for both components of quality because it is strongly correlated with different mechanical and anatomical wood traits [@zobel1989]. Variation in density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. In fact, wood density uniformity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bioenergy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For exemple, a strong interannual variation of the density of the wood leads to a decrease of the mechanical resistance while a relatively uniform density produce wood with better mechanical performance.
 It is therefore important from both a biological point of view and a end-use point of view to have a wood with a uniform density within the stem. \

### Ce que l'on sait de l'uniformité du bois et son importance
In view of the studies cited aboves, we understand that there might be a strong gain at selecting trees that are prone to produce wood of uniform density to get a resistant and higly valuable forest in the next decades. Consequently, studies have been set up to understand the genetic sources of phenotypic variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Mullin2011; @Beaulieu2009]. Hense, it is important to understant where do the variations of density come from in order to better integrated them into genetic selections. In white spruce, there is a typical low frequency radial pattern of variation from pith to bark [@Moore2011]. First a decrease of density during the first years of growth and then an increase from the beginning of the formation of mature wood [@panshin1980]. In addition, this radial variations there are hight frequency interannual variation [@Nicault2010]. Contrary to the radial variation that is under internal control, the interannual variation of the density is influenced mainly by events events to the tree such as the edaphic and climatic conditions [@Nicault2010] that happen at high frequency. \

### Contrôle génétique et environnemental de la densité du bois
Unlike growth that is more controlled by environment than genetic and therefore has low to moderate heritability [@Ivkovich2002a], wood density in white spruce is under strong genetic control with moderate to high heritability [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006] and this makes this trait attractive for genetic improvement. Although variation in wood density is strongly controlled by genetics (30% - 60%) in white spruce, some of the variation is explained by other factors such as growth. Knowing that climate is a determining factor for the growth and physiological processes of trees. As a result, climatic conditions affect wood density. In fact, some climatic factors such as drought can cause stress in trees and cause a variation in wood properties, including density, ring width and conductivity. In addition, long-term growth analyzes have shown that it may increase due to temperature, atmospheric CO2 content and nitrogen deposition [@becker1995; @huang2007; @zaehle2014]. Considering that density can be negatively correlated with growth in spruce [@zhang1995], it is essential to take climat factor into account.

### L'importance de la prise en compte du climat et de l'héritabilité des caractères dans les les programmes de selection génétique
The timeframe of improved forest management being long [@Zobel1972] it is important to anticipate future changes in environmental conditions in these programs [@Gray2016]. This will allow the establishment of plantations capable of adapting to the water stress and possible significant rainfall related to climate change because climate predictions predict more variations in the coming years [@IPCC2007; @pachauri2014]. In fact, with the increasing temperature, the trees are expected to growth faster and thus being of lower quality because of lower density [@corriveau1987; @DeBell1994; @Lindstrom1996; @Dutilleul1998; @Bouriaud2005] hence, the importance of integrating these factors into genetic selection programs. However, to selecting trees that are prone to produce wood of uniform density we must also ensure that these properties will persist in time with climatic change. Thus the selection must also considered the stability of the properties in time or, more accurately, with climate change. Furthermore, heritability of density has been proven in several studies, but there is no evidence to suggest that density variability is hereditary. The heritability of this trait could qualify it for the selection criteria for future breeding programs for producing better quality wood.\

### Le but de notre étude et ce que nous avons fait. 
In this study we quantified and compared heteregoneity of wood density from trees of known parents in order to figure out if some individus produce more uniform wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the homogeneity. By comparing two growth sites we tried to characterize the environmental contribution to homogeity. We ultimalty aim to select and reproduce trees families of interest that are both low-sensitivity to climate change and genetically prone to make homogenous wood. Thus, the most homogeneous families will be identified for reforestation and identify variables to consider in breeding of growing trees with less variable wood.

@@

![Graphic of description of different parameters $b_0$, $b_1$ and $b_2$ of the model. Increasing the value of the parameters $b_0$ and $b_1$ results in an increase in the density at the level of all the rings. The increase in the parameter $b_2$ causes a sharp rise in density after the decrease in the first rings.](graph_table/parameters.pdf)

>>>>

Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of Canadian forest are harvested each year and an additional loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain resource constant, it is therefore essential to compensate the sampling and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better structural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use. \

Improvement programs by genetic selection are in place for the most important commercial species [@DRF2000; @Mullin2011] including white spruce (Picea glauca (Moench) Voss) which has a significant economic value because of its geographical distribution and abundance [@Eyre1980; @Abrahamson2015]. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@Nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly across almost all forested regions which represents a wide range of soils types and a climatic conditions [@Canadian2009]. \

During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood quality has a biologic and an economic component. Biologically, the resistance to mortality by vigour is important to ensure to harvest leaving trees. Economically the value of the wood for end-use is important for industries. Several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. More specifically the homogeneity of the density within the stem has an high importance. I has been demonstrated that trees with a more uniform wood density within the stem are also the more resistant to mortality [@Hietz2013; @osazuwa2014] because a uniform density gives a greater mechanical resistance to the wind and therefore a lower probability of mortality during a climatic event such as a storm. With regard to the economical context, variation of density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. In fact, wood density uniformity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bio energy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For example, a strong inter-annual variation of the density of the wood leads to a decrease of the mechanical resistance while a relatively uniform density produce wood with better mechanical performance [@x,@y]. It is therefore important from both a biological and an end-use points of view to have a wood with a uniform density within the stem. \

In view of the studies cited above, we understand that there might be a strong gain at selecting trees that are prone to produce wood with an homogeneous density to get a resistant and highly valuable forest in the next decades. Consequently, studies have been set up to understand the genetic sources of phenotype variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Mullin2011; @Beaulieu2009]. Hence, it is important to understand where do the variations of density come from in order to better integrated them into genetic selections. In white spruce, there is a typical low frequency radial pattern of variation from pith to bark [@Moore2011] with a density that decreases during the first years of growth and then an increasing from the beginning of the formation of mature wood [@panshin1980]. These variations are under genetic control and are consequently moderately to highly heritable [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006]. This makes this trait attractive for genetic improvement. In addition to these low-frequency variations, there are high frequency inter-annual variations [@Nicault2010] that are influenced mainly by events to the tree such as the edaphic and climatic conditions [@Nicault2010] that happen at high frequency. These variations are not under genetic control but under climate control meaning that climatic conditions also affect wood density. Some climatic factors such as drought can cause stress in trees and cause a variation in wood properties, including density, ring width and conductivity. Long-term growth analyses have shown that density may even increases with the temperature, the atmospheric CO~2~ content and the nitrogen deposition [@becker1995; @huang2007; @zaehle2014]. Wood density can also be negatively correlated with growth. [@zhang1995]. Therefore, with the aim of producing wood of homogeneous density it is essential to take climate factors into account in addition to genetic.

### L'importance de la prise en compte du climat et de l'héritabilité des caractères dans les les programmes de selection génétique
<!-- Ca part en couille il n'y a plus de logique. Quel message veux tu faire passer ? Tu te tires une balle dans le pied en disant l'inverse de que tu as dit avant ! --> 
The time-frame of improved forest management being long [@Zobel1972] it is important to anticipate future changes in environmental conditions in these programs [@Gray2016]. This will allow the establishment of plantations capable of adapting to the water stress and possible significant rainfall related to climate change because climate predictions predict more variations in the coming years [@IPCC2007; @pachauri2014]. In fact, with the increasing temperature, the trees are expected to growth faster and thus being of lower quality because of lower density [@corriveau1987; @DeBell1994; @Lindstrom1996; @Dutilleul1998; @Bouriaud2005] hence, the importance of integrating these factors into genetic selection programs. However, to select trees that are prone to produce wood of homogeneous density we must also ensure that these properties will persist in time with climatic change. Thus the selection must also considered the stability of the properties in time or, more accurately, with climate change. Furthermore, heritability of density has been proven in several studies, but there is no evidence to suggest that density variability is hereditary. The heritability of this trait could qualify it for the selection criteria for future breeding programs for producing better quality wood.\

### Le but de notre étude et ce que nous avons fait. 
In this study we quantified and compared heterogeneity of wood density from trees of known parents in order to figure out if some individuals produce more heterogeneous wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the heterogeneity by comparing two growth sites we tried to characterize the environmental contribution to homogeneity. We ultimately aim to select and reproduce trees families of interest that are both low-sensitivity to climate change and genetically prone to make homogeneous wood. Thus, the most homogeneous families will be identified for reforestation and identify variables to consider in breeding of growing trees with less variable wood.

@@

![Graphic of description of different parameters $b_0$, $b_1$ and $b_2$ of the model. Increasing the value of the parameters $b_0$ and $b_1$ results in an increase in the density at the level of all the rings. The increase in the parameter $b_2$ causes a sharp rise in density after the decrease in the first rings.](graph_table/parameters.png)
